### 2023-01-14 preface/preface.tex

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{preface}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
(Optional, if used placed on a right page next to an empty left page)

I would like to dedicate this work to...
\newpage

\end{preface}