### 2023-02-05 preface/preface.tex

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{preface}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
I would like to dedicate this work to...
\newpage

\end{preface}