### 2023-02-09 MasterThesis.tex
Updates from Overleaf

\documentclass[a4paper,12pt,twoside]{report}

\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry} %Margins
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{setspace}
\usepackage{tocloft}
\usepackage{amsmath}
\usepackage{chngcntr }
\usepackage[toc,page]{appendix}
\usepackage[T1]{fontenc}
\usepackage[nottoc]{tocbibind}
\usepackage[compact]{titlesec}
\usepackage{float}

\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, latexsym}
\usepackage{sidecap}
%---tikz---
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{fadings}

\titlespacing{\section}{0pt}{1ex}{0ex}
\titlespacing{\subsection}{0pt}{1ex}{0ex}
\titlespacing{\subsubsection}{0pt}{1ex}{0ex}

\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}

\include{thesis.preamble}

\begin{document}

\newgeometry{left=2cm,right=2cm} %Only for the title new margins

%Title parameters
\title{Capturing underlying musical structure: an audio-based approach using deep learning}
\author{\href{oriol.colome.font@epidemicsound.com}{Oriol Colomé Font}}
\submitdate{July 2023}
\supervisor{Carlos Lordelo and Carl Thomé}
\cosupervisor{Frederic Font}

\hypersetup{
pdftitle={Your PDF title},
pdfsubject={Your PDF subject},
pdfauthor={Oriol Colomé Font},
pdfkeywords={Music; 
Music Theory; 
Music Information Retrieval; 
Music Structure Analysis; 
Machine Learning; 
Neural Networks; 
Deep Learning}
}

\maketitle

%\maketitle
\restoregeometry

\preface
\cleardoublepage 
\input{dedication/dedication}
\input{acknowledgements/acknowledgements}
\input{preface/preface}
\input{abstract/abstract}

\body

%Introduction of the project
\input{introduction/introduction}
%Music analysis explanation
\input{music analysis/MSA}
%Methods
\input{methods/methods}
%Results
\input{results/results}
%Discussion
\input{discussion/discussion}

\listoffigures
\newpage
\listoftables

% appendices come here
\bibliographystyle{naturemag}
\bibliography{bibliography}

\input{appendix/appendix}

\end{document}