### 2023-03-03 introduction/introduction.tex

\normallinespacing

\input{introduction/Subsections/Introduction}
\input{introduction/Subsections/Motivation}
\input{introduction/Subsections/Objectives}
\input{introduction/Subsections/Structure of the report}

\newpage