### 2023-03-28 methods/methods.tex
Updates from Overleaf

\chapter{Methods}

\input{methods/subsections/machine learning}
\input{methods/subsections/model architecture and training strategy}
\input{methods/subsections/experiment setup and hyperparameter tuning}

\newpage