### 2023-03-28 dedication/dedication.tex
Updates from Overleaf

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{dedication}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
I dedicate this master's thesis to my closest family (Joaquim Colomé, Imma Font, Maria Colomé, and Ariadna Balasch), who has been my constant support and inspiration throughout this journey. Their unwavering love and encouragement have motivated me to pursue my academic goals, and their belief in me has never faltered. I am grateful for their patience, understanding, and sacrifices as I pursued my studies. Their presence in my life has made this journey meaningful, and this thesis is a tribute to their support.

Carl and Carlos
\newpage

\end{dedication}