### 2023-03-30 acknowledgements/acknowledgements.tex
Updates from Overleaf

%\addcontentsline{toc}{chapter}{Acknowledgement}

\begin{acknowledgement}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)

I want to express my sincerest gratitude to...

\begin{itemize}
\item Carlos Lordelo, and Carl Thomé for their invaluable guidance and support throughout my thesis. Your expertise in MIR, ML, and DSP helped shape the direction of my research and make it a success. Your unwavering support and encouragement helped me overcome any difficulties I encountered. I also appreciate the time and effort you invested in mentoring me. Your dedication to your students is truly inspiring, and I am honored to have had the opportunity to work with you.
\vspace*{3mm}
\item Cody Hesse and Sebastian Löf
\vspace*{3mm}
\item Epidemic Sound AB (\textit{Aktiebolag})
\vspace*{3mm}
\item My co-supervisor
\vspace*{3mm}
\item The Music Technology Group (MTG)
\vspace*{3mm}
\item My colleague at My Sheet Music Transcriptions, Oriol López, for his exceptional support and flexibility throughout my thesis. His understanding and willingness to accommodate my academic commitments effectively allowed me to balance my work and research responsibilities. His guidance and advice on professional matters have been instrumental in my growth as a professional. I am genuinely grateful for his unwavering support and encouragement and for being a great mentor.
\vspace*{3mm}
Thank you all for being excellent advisors and playing a crucial role in completing my thesis. I sincerely appreciate everything you have done for me and will always cherish the knowledge and experience I have gained under your guidance.
\vspace*{3mm}
\end{itemize}

\newpage
\end{acknowledgement}