### 2023-04-06 methods/subsections/experiment setup and hyperparameter tuning.tex
Updates from Overleaf

\section{Experiment setup, hyperparameters, and training fine-tuning}

For finetuning and improving the model's behavior, we followed \href{http://karpathy.github.io/2019/04/25/recipe/}{\textit{A Recipe for Training Neural Networks}} \cite{AndrejKarpathyHttp://karpathy.github.io/2019/04/25/recipe/} by \href{https://karpathy.ai/}{Andrej Karpathy} \cite{AndrejKarpathyHttps://karpathy.ai/}. A blog post that discusses a recipe for training neural networks. The author of the blog post tweeted about common mistakes people make when training neural nets and the post aims to expand on those mistakes by explaining how to avoid them through a particular process. Andrej Karpathy notes that there is often a gap between understanding the technical aspects of a neural network and achieving state-of-the-art results.