### 2023-04-13 figures/neural networks/Network_graph.tex

\begin{SCfigure}[12][ht]
	\centering
    \begin{tikzpicture}[shorten >=1pt]
        \tikzstyle{unit}=[draw,shape=circle,minimum size=1.15cm]

        \node[unit](x0) at (0,3.5){$x_0$};
        \node[unit](x1) at (0,2){$x_1$};
        \node(dots) at (0,1){\vdots};
        \node[unit](xd) at (0,0){$x_D$};

        \node[unit](y1) at (4,2.5){$y_1$};
        \node(dots) at (4,1.5){\vdots};
        \node[unit](yc) at (4,0.5){$y_C$};

        \draw[->] (x0) -- (y1);
        \draw[->] (x0) -- (yc);

        \draw[->] (x1) -- (y1);
        \draw[->] (x1) -- (yc);

        \draw[->] (xd) -- (y1);
        \draw[->] (xd) -- (yc);

        \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (-0.5,4) -- (0.75,4) node [black,midway,yshift=+0.6cm]{input layer};
        \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (3.5,3) -- (4.75,3) node [black,midway,yshift=+0.6cm]{output layer};
    \end{tikzpicture}
    \caption[Network graph of a perceptron with $D$ input units and $C$ output units.]{\small{The perceptron consists of $D$ input units and $C$ output units. All units are labeled according to their output: $y_i = f(z_i)$ in the case of output units; $x_i$ in the case of input units. The input values $x_i$ are propagated to each output unit using the weighted sum propagation rule. The additional input value $x_0 := 1$ is used to include the biases as weights.}}
    \label{fig:perceptron}
\end{SCfigure}