### 2023-04-16 figures/my model/SampleCNN.tex
Updates from Overleaf

\begin{tikzpicture}[
  node distance=0.5cm,
  block/.style={draw, rectangle, fill=blue!20, text width=1.5cm, align=center, minimum height=1cm},
  arrow/.style={->, >=stealth, line width=0.5mm}
]

% Input
\node (input) [block] {Input};

% Conv 1
\node (conv1) [block, right=of input] {Conv1D\\128};
\draw[arrow] (input) -- (conv1);

% BatchNorm 1 and ReLU 1
\node (bn1) [block, right=of conv1] {BatchNorm\\ReLU};
\draw[arrow] (conv1) -- (bn1);

% Hidden Layers
\node (hidden1) [block, right=of bn1] {Hidden Layer 1};
\draw[arrow] (bn1) -- (hidden1);

\node (hidden2) [block, right=of hidden1] {Hidden Layer 2};
\draw[arrow] (hidden1) -- (hidden2);

\node (hidden3) [block, right=of hidden2] {Hidden Layer 3};
\draw[arrow] (hidden2) -- (hidden3);

\node (dots) [right=of hidden3] {$\cdots$};
\draw[arrow] (hidden3) -- (dots);

% Conv N
\node (convN) [block, right=of dots] {Conv1D\\512};
\draw[arrow] (dots) -- (convN);

% BatchNorm N and ReLU N
\node (bnN) [block, right=of convN] {BatchNorm\\ReLU};
\draw[arrow] (convN) -- (bnN);

% Average Pooling
\node (avgpool) [block, right=of bnN] {Avg Pool};
\draw[arrow] (bnN) -- (avgpool);

% Dropout (optional)
\node (dropout) [block, right=of avgpool, dashed] {Dropout (optional)};
\draw[arrow, dashed] (avgpool) -- (dropout);

% Fully Connected
\node (fc) [block, right=of dropout] {Linear\\Out\_dim};
\draw[arrow] (dropout) -- (fc);

\end{tikzpicture}