### 2023-04-19 methods/subsections/experiment setup and hyperparameter tuning.tex
Updates from Overleaf

\section{Experiment setup, hyperparameters, and training fine-tuning}

Musically motivated NN \cite{musicmotivCNN}.

For finetuning and improving the model's behavior, we followed \href{http://karpathy.github.io/2019/04/25/recipe/}{\textit{A Recipe for Training Neural Networks}} \cite{AndrejKarpathyHttp://karpathy.github.io/2019/04/25/recipe/} by Andrej Karpathy. A blog post that discusses a recipe for training neural networks. The author of the blog aims to expand on common mistakes people make when training neural nets by explaining how to avoid them through a particular process. Andrej Karpathy notes that there is often a gap between understanding the technical aspects of a neural network and achieving state-of-the-art results.