### 2023-05-03 introduction/Subsections/Structure of the report.tex
Updates from Overleaf

\section{Structure of the Report}

In this report, we have organized the information into several sections and sub-sections to provide a clear and comprehensive overview of the topic. The structure of the report is as follows:

\begin{itemize}
\item Introduction
  \begin{itemize}
  \item Motivation
  \item About music aural skills and high-level perceptual concepts
  \item Objectives
  \item Assumptions
  \item Structure of the report
  \end{itemize}
\item Methods
  \begin{itemize}
  \item Introduction
  \item Artificial Neural Networks (ANNs)
  \item About ANNs
  \item Deep Learning (DL)
  \item Convolutional Neural Networks (CNNs)
  \item Model architecture and training strategy
  \item Contrastive Learning of Musical Representations (CLMR)
  \item Encoder
  \item Loss function
  \item Experiment setup, hyperparameters, and training fine-tuning
  \end{itemize}
\item Evaluation
  \begin{itemize}
  \item Beyond Digital Signal Processing (DSP)
  \item Sample MIR task
  \end{itemize}
\item Results
  \begin{itemize}
  \item Summary
  \item Recommendations
  \end{itemize}
\item Discussion
  \begin{itemize}
  \item Discussion
  \item Conclusion
  \end{itemize}
\item Future work
  \begin{itemize}
  \item Summary
  \end{itemize}

\item References
  \begin{itemize}
  \item Sources Cited
  \end{itemize}
\end{itemize}