### 2023-05-05 figures/neural networks/Single_CNN.tex
Updates from Overleaf

\begin{SCfigure}[2\sidecaptionrelwidth][ht]
	\centering
        \scalebox{0.8}{
	\begin{tikzpicture}
		\node at (1.5,4){\begin{tabular}{c}input data or \\ feature map\end{tabular}};

		\draw (0,0) -- (3,0) -- (3,3) -- (0,3) -- (0,0);

		\draw (2,2) -- (2.5,2) -- (2.5,2.5) -- (2,2.5) -- (2,2);
		\draw (2,0.5) -- (2.5,0.5) -- (2.5,1) -- (2,1) -- (2,0.5);
		\draw (1,1) -- (1.5,1) -- (1.5,1.5) -- (1,1.5) -- (1,1);

		\draw (2.5,2) -- (7,3.25);
		\draw (2.5,2.5) -- (7,3.25);

		\draw (2.5,1) -- (5.75,0.25);
		\draw (2.5,0.5) -- (5.75,0.25);

		\draw (1.5,1.5) -- (5.5,1.25);
		\draw (1.5,1) -- (5.5,1.25);

		\node at (5.75,4){\begin{tabular}{c}output feature maps\end{tabular}};

		\draw[fill=black,opacity=0.2,draw=black] (5.5,1.5) -- (7.5,1.5) -- (7.5,3.5) -- (5.5,3.5) -- (5.5,1.5);
		\draw[fill=black,opacity=0.2,draw=black] (5,1) -- (7,1) -- (7,3) -- (5,3) -- (5,1);
		\draw[fill=black,opacity=0.2,draw=black] (4.5,0.5) -- (6.5,0.5) -- (6.5,2.5) -- (4.5,2.5) -- (4.5,0.5);
		\draw[fill=black,opacity=0.2,draw=black] (4,0) -- (6,0) -- (6,2) -- (4,2) -- (4,0);
	\end{tikzpicture}
        }
	\caption[Illustration of a 2D convolutional layer.]{\small{Illustration of a 2D single convolutional layer. If layer $l$ is a convolutional layer, the input data or a feature map of the previous layer is convolved by different kernels to yield the output feature maps of layer $l$.}}
	\label{fig: convolutional layer}
\end{SCfigure}