### 2023-05-13 evaluation/subsections/beyond dsp.tex
Updates from Overleaf

\section{Music Structure Analysis as downstream task}

Let's pick a core MIR task as a sample to display such complexity: Music Structure Analysis (MSA), an interdisciplinary field that aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity. The checkerboard kernel technique is a simple and effective method for MSA based on the homogeneity principle. The kernel, with a checkerboard-like structure, is convolved over the main diagonal of a Self-Similarity Matrix (SSM) such as $S_{ij}$ is the similarity between time frames $i$ and $j$, $\vec{x}_i$ is the vector representation of time frame $i$, and $\left| \vec{x}_i \right|$ is the Euclidean norm of vector $\vec{x}_i$. The dot product of two vectors $\vec{x}_i$ and $\vec{x}_j$ is denoted by $\vec{x}_i \cdot \vec{x}_j$.

\input{introduction/subsections/equations/SSM equation}

This yields a novelty curve highlighting sudden changes in the selected musical features from which to extract the segment boundaries.

\input{introduction/subsections/equations/Novelty equation}

Mathematical operations are used on predefined principles such as homogeneity and repetition to analyze a musical structure. It explores features extracted from a musical signal and identifies peaks in a novelty curve to extract segment boundaries. In contrast, human aural skills rely on the perception and interpretation of music by human listeners based on cultural and academic background. They can capture subtle nuances of musical structure, but those still are subjective and can vary across listeners.

Combining technical tasks, aural skills, and music perception is challenging due to the complexity and subjectivity of musical perception, discrepancies between the features extracted by DSP techniques and the features perceived by humans, and the variability of musical signals and individual differences in musical perception.

Purely mathematical approaches may not always reflect how humans perceive music and may not capture the subtle variations in timing or harmonic relationships critical to musical perception. Variability in musical signals due to instrumentation, genre, and performance style further complicates the development of these techniques that can be generalized to different musical contexts.

\input{figures/neural networks/r128 sphere}

\section{SALAMI dataset}
\subsection{Overview}

SALAMI (Structural Analysis of Large Amounts of Music Information) \cite{Smith2011DESIGNANNOTATIONS} aims to conduct massive structural analyses of different types of music. The project covers diverse music, from Western pop to Indian classical and from live to studio recordings.

\subsubsection{Formal Analysis Definition}
Formal analysis, in this context, doesn't refer to classifying the piece into a formal type (e.g., sonata, song, or canon) or reducing it into its Ursatz as in Schenkerian analysis. Instead, it means organizing and dividing the piece into specific sections and understanding how they relate.

%%%%%%%%%%

\subsubsection{Approaches to Music Analysis}

The method proposed combines aspects of the following three approaches, separating the organization of instrumentation, musical material, and formal function. This allows for the analysis of a wide variety of music pieces in a consistent way using a constrained vocabulary. It's important to note that music structure is often hierarchical, and the appropriate timescale for analysis can be hard to determine. The proposed method includes a few markers to partly address this issue.

Perceptual Approach: This involves listening for prominent harmonic or rhythmic boundaries in the piece to segment and apply labels to sections based on their similarity. However, this approach might not reflect the different structural roles of similar-sounding sections.

Functional Approach: This involves dividing the song into its functional parts, like verses, choruses, intros, outros, etc. This might result in the same music being labeled differently based on its function or different musical ideas having the same functional label.

Transcription Approach: This involves recording musical parameters like chord patterns and instrumentation directly. While this method may be less subjective, it may not provide a comprehensive view of the piece's structure.