### 2023-05-15 methods/subsections/machine learning.tex

\section{Introduction}

In this chapter, we present the methodology employed in this study.

We provide an overview of the Machine Learning (ML) and Deep Learning techniques that have played a pivotal role in advancing the MIR field. We then discuss the importance of feature extraction and selection in MIR and how automatic feature learning has facilitated the development of more robust models.

Subsequently, we delve into the concept of musically motivated neural network architectures, which incorporate domain-specific knowledge and musical ideas into the design of neural networks. We explore techniques proposed in the literature for blending musical concepts and perceptual approaches.

We then outline the specific MIR tasks that our study focuses on and discuss the evaluation metrics used to assess the performance of the proposed models. Finally, we present the experimental setup, including the datasets used for training and validation, the preprocessing steps, and the training parameters.

Throughout this chapter, we aim to provide a comprehensive understanding of the methodology employed in our research, laying the foundation for the presentation and discussion of the results in the following chapters.

\section{Artificial Neural Networks (ANNs)}
A neural network, formally known as an Artificial Neural Network (ANN), is a computational model inspired by the structure and functionality of biological neural networks. It is a mathematical framework designed to process information and perform tasks like pattern recognition, classification, and regression by mimicking how biological neurons and synapses operate. ANNs comprise interconnected layers of artificial neurons or nodes, each performing a mathematical transformation on the input data to generate an output. The connections between these neurons, often called "weights," are adjusted during learning to optimize the network's performance on a given task.

\subsection{About ANNs}
The concept of neural networks can be traced back to the mid-20th century, with the pioneering work of Warren McCulloch and Walter Pitts in 1943. They proposed a mathematical model of a neuron known as the McCulloch-Pitts neuron, which provided a foundation for future research in artificial intelligence. However, Frank Rosenblatt's invention of the perceptron in 1957 marked the beginning of modern neural networks. The perceptron is a single-layer neural network capable of learning linearly separable patterns.

%Single neuron
\input{figures/neural networks/Single Neuron}

%NN
\input{figures/neural networks/Network_graph}

In a more technical description, an ANN consists of an input layer, one or more hidden layers, and an output layer. Each neuron in a layer receives input from the previous layer and computes an output value using an activation function. The output of the last layer represents the prediction or classification result.

%Activation equation
\input{figures/neural networks/activation}

%Multi-layer_NN
\input{figures/neural networks/Multi-layer_NN}

\section{Deep Learning (DL)}

Deep learning is a subfield of machine learning that employs deep neural networks to learn hierarchical data representations, enabling computational models with multiple processing layers to learn data abstractions across numerous levels. These methods have dramatically advanced the state-of-the-art in various domains, including speech recognition, visual object recognition, object detection, drug discovery, genomics, and more \cite{LeCun2015DeepLearning}. The effectiveness of deep learning arises from its nonlinear modeling capabilities and scalability with large datasets.

A central algorithm in deep learning is backpropagation, which uses the forward and backward pass process to efficiently compute gradients of the loss function concerning the network's parameters. The forward pass involves propagating input data through the network to generate predictions. In contrast, the backward pass calculates the error between predictions and ground truth and subsequently returns it through the network. This procedure updates the network's parameters using gradient descent, minimizing the loss function and enhancing the model's performance.

Deep learning has led to breakthroughs in processing complex, high-dimensional data, including images, video, speech, and audio, primarily due to deep convolutional nets. Simultaneously, recurrent nets have illuminated sequential data analysis such as text and speech. Deep learning discovers intricate structures in large datasets, improving our understanding and processing capabilities across various applications.

%Backpropagation
\input{figures/neural networks/Back_propagation}

\newpage