### 2023-05-17 figures/neural networks/Single Neuron.tex
Updates from Overleaf

\begin{SCfigure}[2\sidecaptionrelwidth][h]
	\centering
	\begin{tikzpicture}[shorten >=1pt,->]
		\tikzstyle{unit}=[draw,shape=circle,minimum size=1.15cm]

		\node[unit](p) at (2,1){$y$};
		\node(dots) at (-0.25,1){\vdots};

		\draw (0,2.5) node[xshift=-10]{$w_0$} -- (p);
		\draw (0,1.75) node[xshift=-10]{$x_1$} --(p);
		\draw (0,0) node[xshift=-10]{$x_D$} -- (p);
		\draw (p) -- (3,1) node[xshift=30]{$y := a(z)$};
	\end{tikzpicture}
	\caption[Single processing units and its components.]{\small{Single processing unit and its components. The activation function is denoted by $a$ and applied to the actual input $z$ of the unit to form its output $y = a(z)$. $x_1, \ldots, x_D$ represent input from other units within the network; $w_0$ represents the bias, external input to the unit. The propagation rule maps all inputs onto the actual input $z$.}}
	\label{fig:processing-unit}
\end{SCfigure}