### 2023-05-18 acknowledgements/acknowledgements.tex
Updates from Overleaf

%\addcontentsline{toc}{chapter}{Acknowledgement}

\begin{acknowledgement}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)

I want to extend my deepest gratitude to...

\begin{itemize}
\item Carlos Lordelo and Carl Thomé for their invaluable guidance and support throughout my thesis. Their MIR, ML, and DSP expertise has been instrumental in shaping my research and driving it toward success. I am truly grateful for their unwavering support and encouragement, which helped me surmount any difficulties I encountered during this journey. Their time, effort, and dedication in mentoring me have not gone unnoticed. They inspire me with their commitment to their students, and I feel honored to have had the opportunity to work with them.

Words fall short of my profound gratitude for their consistent, knowledgeable, and steadfast support. I am beyond grateful that their dedication has left an indelible mark on my academic journey.

\vspace*{3mm}
\item Cody Hesse and Sebastian Löf
\vspace*{3mm}
\item Epidemic Sound AB (\textit{Aktiebolag})
\vspace*{3mm}
\item My co-supervisor
\vspace*{3mm}
\item The Music Technology Group (MTG)
\vspace*{3mm}
\item My colleague at My Sheet Music Transcriptions, Oriol López, for his exceptional support and flexibility throughout my thesis. His understanding and willingness to accommodate my academic commitments effectively allowed me to balance my work and research responsibilities. His guidance and advice on professional matters have been instrumental in my growth as a professional. I am genuinely grateful for his unwavering support and encouragement and for being a great mentor.
\vspace*{3mm}
Thank you all for being excellent advisors and playing a crucial role in completing my thesis. I sincerely appreciate everything you have done for me and will always cherish the knowledge and experience I have gained under your guidance.
\vspace*{3mm}
\end{itemize}

\newpage
\end{acknowledgement}