### 2023-05-25 methods/methods.tex
Updates from Overleaf

\chapter{Methods}

\input{methods/subsections/intro}
\input{methods/subsections/deep learning}
\input{methods/subsections/implementation}

\newpage