### 2023-05-25 methods/subsections/intro.tex

\section{Introduction}

Throughout this chapter, we furnish an exhaustive understanding of the methodical process employed in our research, thereby setting the stage for the presentation and discussion of the results in subsequent chapters.

Our approach could be broadly categorized as cognitive modeling, as it strives to mimic human cognitive processes and problem-solving through a computerized model. We champion exposure-based learning in music and active engagement with an array of musical styles, genres, techniques, and learning methods to foster well-rounded musical proficiency and efficient performance with novel data in practical applications.