### 2023-06-15 results/results.tex

\chapter{Results}

\input{results/subsections/tables}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/metrics.png}
    \caption{Caption}
    \label{fig:boxplotmetrics}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/boudaryfscore.png}
    \caption{Caption}
    \label{fig:boxplotfmeasure}
\end{figure}

\newpage