### 2023-06-29 results/results.tex
Updates from Overleaf

\chapter{Results}

\input{results/subsections/tables}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/metrics.png}
    \caption[Metric comparison for different audio features. Boxplot]{Boxplot visual comparison of precision, recall, and f-measure for different features obtained using the SF algorithm \cite{sf} on the SALAMI dataset. Sliding windowed segments across the audio signal input signal is 4 seconds}
    \label{fig:boxplotmetrics}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/boudaryfscore.png}
    \caption[F-measure comparison for different audio features. Boxplot]{Boxplot visual comparison of f-measure for different features obtained using the SF algorithm \cite{sf} on the SALAMI dataset. Sliding windowed segments across the audio signal input signal is 4 seconds}
    \label{fig:boxplotfmeasure}
\end{figure}

\newpage