### 2023-06-30 introduction/Subsections/Previous work.tex
Updates from Overleaf

\section{Related work}

Related research \cite{deepfeaturesegment, SalamonDeepSegmentation} explores how the use of audio embeddings learned via unsupervised methods can enhance the performance of music segmentation algorithms leveraging the power of deep learning and data-driven feature learning to improve upon traditional, manually-engineered methods of music segmentation.

Both studies conclude that using deep learning techniques to derive audio embeddings improves the performance of music segmentation algorithms, thus providing a better alternative to traditional manual feature engineering methods.