### 2023-07-02 introduction/Subsections/Objectives.tex

\section{Objectives}

Can deep audio embeddings be learned from reordering, scrambling, and augmenting via audio effects sequences of musical information to improve unsupervised music boundary detection?

This study explores the high-level tonal structure of music by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn and analyze music embeddings that can be applied to subsequent MIR tasks, with a particular focus on music boundary detection.

Our research aims to develop and learn a unified numerical understanding (or embedding \footnote{A learned representation or embedding is a numerical output - usually a fixed-size vector - produced by a machine-learning model. Good representations should be versatile across various tasks and require limited supervision \cite{Turian2022HEAR:Representations}.}) for a piece of music that integrates the high-level relationships between musical elements as they unfold over time.

The research employs MIR computational methodologies and musicological perspectives to explore musical compositions across various styles and genres.

This approach aims to replicate human auditory capabilities by understanding and identifying abstract and semantic musical elements independent of their sonic qualities.