### 2023-07-02 methods/methods.tex

\chapter{Methods}

This work concentrates on employing unsupervised training of neural networks to extract meaningful features for music segmentation methodologies. As assertively stated by \cite{deepfeaturesegment}, this constitutes a logical evolution in the application of contemporary machine learning techniques to the issue of music segmentation.

Our work confronts the persistent challenge of acquiring labeled data for music segmentation, an endeavor often time-consuming and expensive. By capitalizing on the vast volumes of unlabeled music data, we bypass the need for laborious hand annotation, thereby offering a pathway to potentially more comprehensive generalization given the broader dataset.

\input{methods/subsections/deep learning}
\input{methods/subsections/implementation}

\newpage