### 2023-07-04 results/results.tex

\chapter{Results}

\input{results/subsections/tables}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/boxplot.png}
    \caption[Metric comparison for different audio features.]{\small{Boxplot visual comparison of different features' average precision, recall, and f-measure. Sliding windowed segments across the audio signal input signal is 4 seconds.}}
    \label{fig:boxplotmetrics}
\end{figure}

\newpage