### 2023-07-04 results/subsections/tables.tex

As stated in the previous section, the evaluation results were obtained using standard and commonly used MIR tools, frameworks, and metrics. These include the MSAF \cite{MSAF} implementation for segmentation algorithms, the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS} for evaluation ground truth, and $mireval$ Python package \cite{RaffelMir_eval:METRICS} for metric computation.

The GitHub repository containing all the code needed to run the experiments can be found \href{https://github.com/oriolcolomefont/Master-Thesis.git}{HERE}.The results are displayed in the table \ref{ta:results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}X>{\centering\arraybackslash}X>{\centering\arraybackslash}X}
\toprule
\textbf{Training dataset [Ref]}  & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\midrule
\addlinespace
GTZAN \cite{GTZAN} & 0.228 & 0.171 & 0.185 \\
\addlinespace
MSD \cite{MSD} &  \textbf{0.333} &  \textbf{0.280} & \textbf{0.288} \\
\addlinespace
\bottomrule
\end{tabularx}
\caption[Caption]{\small{Comparison of precision, recall, and f-measure for GTZAN-trained versus MSD-trained embeddiograms on SALAMI dataset computed using the Structural Feature (SF)\cite{sf} algorithm.}}
\label{tab:comparison_table}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newsavebox\embeddiobSF
\begin{lrbox}{\embeddiobSF}
   $\begin{aligned}
     \textbf{0.333} & \quad 0.280 & \quad 0.288
    \end{aligned} $
\end{lrbox}

\newsavebox\embeddiobFoote
\begin{lrbox}{\embeddiobFoote}
   $\begin{aligned}
     0.275 & \quad 0.318 & \quad 0.280
    \end{aligned} $
\end{lrbox}

\newsavebox\embeddiobCNMF
\begin{lrbox}{\embeddiobCNMF}
   $\begin{aligned}
     0.248 & \quad 0.296 & \quad 0.254
    \end{aligned} $
\end{lrbox}

\newsavebox\pcpSF
\begin{lrbox}{\pcpSF}
   $\begin{aligned}
     0.311 & \quad 0.324 & \quad 0.305
    \end{aligned} $
\end{lrbox}

\newsavebox\pcpFoote
\begin{lrbox}{\pcpFoote}
   $\begin{aligned}
     0.288 & \quad \textbf{0.331} & \quad 0.295
    \end{aligned} $
\end{lrbox}

\newsavebox\pcpCNMF
\begin{lrbox}{\pcpCNMF}
   $\begin{aligned}
     0.228 & \quad 0.310 & \quad 0.250
    \end{aligned} $
\end{lrbox}

\newsavebox\tonnetzSF
\begin{lrbox}{\tonnetzSF}
   $\begin{aligned}
     0.312 & \quad 0.312 & \quad 0.300
    \end{aligned} $
\end{lrbox}

\newsavebox\tonnetzFoote
\begin{lrbox}{\tonnetzFoote}
   $\begin{aligned}
     0.272 & \quad 0.317 & \quad 0.280
    \end{aligned} $
\end{lrbox}

\newsavebox\tonnetzCNMF
\begin{lrbox}{\tonnetzCNMF}
   $\begin{aligned}
     0.212 & \quad 0.306 & \quad 0.237
    \end{aligned} $
\end{lrbox}

\newsavebox\cqtSF
\begin{lrbox}{\cqtSF}
   $\begin{aligned}
     0.311 & \quad \textbf{0.339} & \quad \textbf{0.312}
    \end{aligned} $
\end{lrbox}

\newsavebox\cqtFoote
\begin{lrbox}{\cqtFoote}
   $\begin{aligned}
     \textbf{0.570} & \quad 0.311 & \quad \textbf{0.353}
    \end{aligned} $
\end{lrbox}

\newsavebox\cqtCNMF
\begin{lrbox}{\cqtCNMF}
   $\begin{aligned}
     \textbf{0.296} & \quad \textbf{0.311} & \quad \textbf{0.287}
    \end{aligned} $
\end{lrbox}

%%%%%

\begin{table}
  \centering
  \begin{adjustbox}{width=\textwidth}
  \begin{threeparttable}
    \begin{tabular}{c|c|c|c|c} 
\toprule
     \textbf{Feature} & \textbf{SF} & \textbf{Foote} & \textbf{CNMF} 
     \\ \midrule 
     PCP & \usebox{\pcpSF} & \usebox{\pcpFoote} & \usebox{\pcpCNMF} \\\hline
     Tonnetz & \usebox{\tonnetzSF} & \usebox{\tonnetzFoote} & \usebox{\tonnetzCNMF} \\\hline
     CQT & \usebox{\cqtSF} & \usebox{\cqtFoote} & \usebox{\cqtCNMF} \\\hline
     Embeddiogram & \usebox{\embeddiobSF} & \usebox{\embeddiobFoote} & \usebox{\embeddiobCNMF} \\
     \bottomrule
    \end{tabular}
    \caption[Metric comparison: audio features and segmentation algorithms]{Comparison of precision (left column), recall (middle column), and f-measure (right column) metrics for different features using the Structural Feature (SF)\cite{sf}, Checkerboard-like Kernel (Foote) \cite{Foote2000AutomaticNovelty}, and Convex Non-negative Matrix Factorization (CNMF) \cite{NietoCONVEXIDENTIFICATION} algorithms on the SALAMI dataset. The sliding windowed segments across the input signal is 4 seconds long.}\label{ta:results}
  \end{threeparttable}
  \end{adjustbox}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}p{4.5cm}XXXX}
\toprule
\thead{\textbf{Authors [Ref], Year}} & \thead{\textbf{Input}} & \thead{\textbf{Method}} & \thead{\textbf{F-measure}} \\
\midrule
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & CQT Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
\cite{deepfeaturesegment}, 2019 & CQT & CNN  & \textbf{0.535} \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}
\label{tab:comparison_table}
\end{table}