### 2023-07-06 acknowledgements/acknowledgements.tex
Updates from Overleaf

%\addcontentsline{toc}{chapter}{Acknowledgement}

\begin{acknowledgement}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)

I want to extend my deepest gratitude to...

\begin{itemize}
\item Carl Thomé and Carlos Lordelo, whose expertise in MIR, ML, and DSP has been pivotal to the success of my thesis. Their guidance and encouragement have enabled me to overcome the challenges encountered during this journey. Their commitment to mentorship is truly inspiring, and their influence has significantly shaped my academic path. 
Working under their mentorship has been both a privilege and an honor.

\vspace*{3mm}
\item Cody Hesse and Sebastian Löf. Their generosity in sharing their time, knowledge, and insights has critically enhanced the quality of this research work.

\vspace*{3mm}
\item The Music Technology Group (MTG) and Universitat Pompeu Fabra (UPF). My profound thanks go to Xavier Serra for his faith in accepting a non-traditional profile into the master's program. I am equally grateful to Pedro Ramoneda and Genís Plaja for their consistent support throughout my journey.

\vspace*{3mm}

I sincerely thank you all for your integral roles in completing my thesis. I deeply value the knowledge and experience gained under your guidance and will cherish it always.

\vspace*{3mm}
\end{itemize}

\newpage
\end{acknowledgement}