### 2023-07-06 dedication/dedication.tex

%\addcontentsline{toc}{chapter}{Acknowledgement}
\begin{dedication}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
I dedicate this master's thesis to my closest family and friends, who have supported and inspired me throughout this journey. Their love and encouragement have motivated me to pursue my academic goals, and their belief in me has never faltered. I am grateful for their patience, understanding, and sacrifices as I pursued my studies.

Their presence in my life has made this journey meaningful, and this thesis is a tribute to their support.

\newpage

\end{dedication}