### 2023-07-07 results/results.tex

\chapter{Results}

\input{results/subsections/tables}

\newpage