### 2023-08-05 acknowledgements/acknowledgements.tex

%\addcontentsline{toc}{chapter}{Acknowledgement}

\begin{acknowledgement}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)

I want to extend my deepest gratitude to...

\begin{itemize}
\item Carl Thomé and Carlos Lordelo, supervisors and fellow MIR researchers. Their expertise has been pivotal to the success of this thesis. Their guidance, supervision, and encouragement have enabled me to overcome the challenges encountered during this journey. Their commitment to mentorship is genuinely inspiring, and their influence has significantly shaped my academic path.

Working with them has been an absolute privilege.

\vspace*{3mm}
\item Cody Hesse and Sebastian Löf, fellow interns. Their generosity in sharing their time, knowledge, and insights has critically enhanced the quality of this research work.

\vspace*{3mm}
\item The Music Technology Group (MTG) and Universitat Pompeu Fabra (UPF). My profound thanks go to Xavier Serra for his faith in accepting a non-traditional profile into the master's program.

\vspace*{3mm}

I sincerely thank you for your integral roles in completing this thesis. I deeply value the knowledge and experience gained under their guidance and will cherish it always.

\vspace*{3mm}
\end{itemize}

\newpage
\end{acknowledgement}