### 2023-08-21 methods/methods.tex

\chapter{Methods}

This work employs unsupervised training of neural networks to extract meaningful features for music segmentation methodologies. As assertively stated by \cite{deepfeaturesegment}, this constitutes a logical evolution in applying contemporary machine learning techniques to music segmentation.

\input{methods/subsections/deep learning}
\input{methods/subsections/implementation}

\newpage